### 2023-10-12 reservoirnet : an R package for reservoir computing/reservoirnet : an R package for reservoir computing.Rmd
update jss

The input layer $u(t)$ is an $M$-dimension vector which corresponds to the values of the input time series at time $t$ where $t = 1, …, T$. The reservoir layer $x(t)$ is an $Nres$-dimensional vector where $Nres$ is the number of nodes in the reservoir. The value $x(t)$ is defined as follow:

where

The leaking rate alpha define the rate of update of the nodes. The closer alpha is to 1, the higher the reservoir is sensitive to new inputs (i.e u(t)). Therefore, the reservoir state at time t+1 denoted x(t+1) depends on the reservoir state at the previous time (i.e x(t)) and the new inputs (i.e u(t+1)). Both Win and W are random matrices of size Nres x M and Nres x Nres respectively.

Win is a dense matrix generated using a bernouilli distribution where each value can be either -IS(m) or IS(m) with an equal probability where m = 1, …, M corresponds to a given feature in the input layer. The input scaling, denoted IS, is an hyperparameter coefficient wich can be common to all features from the input layer or specific to each feature m. In that case, the more important the feature is, the greater should be its input scaling.  W is a sparse matrix where values are generated base on a gaussian distribution with mean 0 and standard deviation equal to 1. Then the matrix W is rescaled according to the spectral radius, an hyperparameter defining the highest eigen value of W.

@@

where x(t) and y(t) are accumulated in X and Y respectively such that:
and

The parameter lambda is the ridge penalisation which aims to prevent overfitting.

@@

    1. Initialize:  
        - Population parameters : Npop = 200, Ne = 100, Ngeneration = 30
        - Cross-over parameters : Ntournament = 2 
        - Mutation parameters : pmutQuant = .5, pmutCat = .25, sigma = 1 (0.1 for leaking rate as the hyperparameter range is shorter)
        - Define the hyperparameter_list, a list containing each param.
        - For each param in hyperparameter_list define hyperparameter search_space: define the lower and upper bound, the type (numeric, integer, categorical) and if it should be log transformed (True, False)

    2. Define tournamentSelection(Npop, Ntournament): 
        - Get the best Npop individuals from previous generations 
        - Randomly select Ntournament challengers among them
        - Get the best individuals from the challengers
        - Repeat twice to get a father and a mother 
        - return father, mother

    3. Define crossoverMutation(pere.param.value, mere.param.value, param.search_space): 
        - sample alp and mut from a Uniform(0,1) distribution
        - If search_space.log is True then  pere.param.value, mere.param.value = log10(pere.param.value), log10(mere.param.value)
        - Cross-over, and define child.param.value as:
            - Numerical : alp*pere.param.value + (1-alp)*mere.param.value
            - Integer : round( alp*pere.param.value + (1-alp)*mere.param.value)
            - Categorical : random choice (list = [pere.param.value, mere.param.value], prob = [alp, 1-alp]) 
        - Mutation if mut < pmutQuant or mut < pmutCat for quantitative and categorical features respectively and update child.param.value
            - Categorical : Randomly selected an other modality 
            - Integer : Randomly add or substract 1 
            - Numerical : Add sigma * N(0,1) 
        - If search_space.log is True then child.param.value = 10^child.param.value
        - Return child.param.value

    4. Define geneticSearch(Npop, Ntournament):
            - pere, mere = tournamentSelection(Npop, Ntournament)
            - for param in  hyperparameter_list:
                - child.param.value = crossoverMutation(pere.param.value, mere.param.value, param.search_space)
            - return child.param

Step 1: Initialize genetic algorithm

Sample Npop individuals (i.e set of hyperparameters) using random search

Step 2: Run genetic algorithm
generation = 1
While generation <  Ngeneration:	
	generation = generation + 1
	sample Npop individuals using  geneticSearch method

>>>>

The input layer $u(t)$ is an $M$-dimension vector which corresponds to the values of the input time series at time $t$ where $t = 1, …, T$. The reservoir layer $x(t)$ is an $N_{res}$-dimensional vector where $N_{res}$ is the number of nodes in the reservoir. The value $x(t)$ is defined as follow:

$$x( t+1 ) = ( 1 - \alpha )  x ( t) + \alpha \: tanh( W x(t) + W_{in} u(t+1) ) \text{ where } \alpha \in [0, 1 ]$$

The leaking rate alpha define the rate of update of the nodes. The closer $\alpha$ is to $1$, the higher the reservoir is sensitive to new inputs (i.e $u(t)$). Therefore, the reservoir state at time $t+1$ denoted $x(t+1)$ depends on the reservoir state at the previous time (i.e $x(t)$) and the new inputs (i.e $u(t+1)$). Both $W_{in}$ and $W$ are random matrices of size $Nres \times M$ and $Nres \times Nres$ respectively.

$W_{in}$ is a dense matrix generated using a bernouilli distribution where each value can be either $-I_{scale}(m)$ or $I_{scale}(m)$ with an equal probability where $m = 1, …, M$ corresponds to a given feature in the input layer. The input scaling, denoted $I_{scale}$, is an hyperparameter coefficient which can be common to all features from the input layer or specific to each feature $m$. In that case, the more important the feature is, the greater should be its input scaling. $W$ is a sparse matrix where values are generated base on a gaussian distribution $\mathcal{N}(0,1)$. Then the matrix $W$ is rescaled according to the spectral radius, an hyperparameter defining the highest eigen value of $W$.

@@

$$W_{out} = YX^T ( XX^T + \lambda  I)^{ -1 }$$

Where x(t) and y(t) are accumulated in X and Y respectively such that:

$$X = \begin{bmatrix} x(1) \\ x(2) \\ ... \\ x(T) \end{bmatrix}
\text{ and } Y = \begin{bmatrix} y(1) \\ y(2) \\ ... \\ y(T) \end{bmatrix}$$

The parameter $\lambda$ is the ridge penalisation which aims to prevent overfitting.

@@

1. Initialize :
  - Population parameters : Npop = 200, Ne = 100, Ngeneration = 30
  - Cross-over parameters : Ntournament = 2 
  - Mutation parameters : pmutQuant = .5, pmutCat = .25, sigma = 1 (0.1 for leaking rate as the hyperparameter range is shorter)
  - Define the hyperparameter_list, a list containing each param.
  - For each param in hyperparameter_list define hyperparameter search_space : define the lower and upper bound, the type (numeric, integer, categorical) and if it should be log transformed (True, False)

2. Define tournamentSelection(Npop, Ntournament): 
    - Get the best Npop individuals from previous generations 
    - Randomly select Ntournament challengers among them
    - Get the best individuals from the challengers
    - Repeat twice to get a father and a mother 
    - return father, mother 
3. Define crossoverMutation(pere.param.value, mere.param.value, param.search_space): 
    - sample alp and mut from a Uniform(0,1) distribution
    - If search_space.log is True then  pere.param.value, mere.param.value = log10(pere.param.value), log10(mere.param.value)
    - Cross-over, and define child.param.value as :
        - Numerical : alp*pere.param.value + (1-alp)*mere.param.value
        - Integer : round( alp*pere.param.value + (1-alp)*mere.param.value)
        - Categorical : random choice (list = [pere.param.value, mere.param.value], prob = [alp, 1-alp]) 
    - Mutation if mut < pmutQuant or mut < pmutCat for quantitative and categorical features respectively and update child.param.value
        - Categorical : Randomly selected an other modality 
        - Integer : Randomly add or substract 1 
        - Numerical : Add sigma * N(0,1) 
    - If search_space.log is True then child.param.value = 10^child.param.value
    - Return child.param.value

4. Define geneticSearch(Npop, Ntournament) :
    - pere, mere = tournamentSelection(Npop, Ntournament)
    - for param in  hyperparameter_list :
        - child.param.value = crossoverMutation(pere.param.value, mere.param.value, param.search_space)
    - return child.param

Step 1 : Initialize genetic algorithm

- Sample Npop individuals (i.e set of hyperparameters) using random search

Step 2 : Run genetic algorithm
- generation = 1
While generation <  Ngeneration :	
    generation = generation + 1
    sample Npop individuals using geneticSearch method
